IML Term project paper

### 2022-12-06 paper.rmd

```{r load data}
npf <- read.csv("npf_train.csv")
```

@@

To understand the data, we ... (about the visualizations, what to include?)

Many classifiers are affected by correlation and collinearity between variables. As one would expect, we found that many of the variables that are of the same phenomenon are correlated, as are variables such as humidity and temperature. We take a more detailed look into this in the section *Steps we took to select good features and model parameters*.

@@

(The data is over non-consecutive days, ok, but are there any significant gaps e.g. a month?)
(How many unique variables are there if we ignore the column that are just of the same variable at different heights?)

@@

>>>>

```{r load data, echo=FALSE}
npfOrig <- read.csv("npf_train.csv")
# Create class2
npfOrig$class2[npfOrig$class4 == "nonevent"] <- 0
npfOrig$class2[npfOrig$class4 != "nonevent"] <- 1
# Add date as row name
rownames(npfOrig) <- npfOrig[, "date"]
# Remove columns id, date, class4, partlybad
npfData <- npfOrig[, -(1:4)]
```

@@

Many classifiers are affected by correlation and colinearity between variables. As expected, we found that many of the variables describing the same phenomenon are correlated, as are variables such as humidity and temperature. We take a more detailed look into this in the section *Steps we took to select good features and model parameters*.

@@

(The data is over non-consecutive days, ok, but are there any significant gaps e.g. a month?)

We found that there are 38 unique columns if we select measurements from only one height. We ended up selecting height 16.8 meter as some measurements were taken only at that height. Plots below describe the correlations in the original data and correlations after omitting all but the selected height.

### Selecting columns by correlation
```{r original correlation plot, echo=FALSE, fig.dim=c(7, 7), message=FALSE}
library(corrplot)
# Original correlations
cmData <- cor(npfData)
corrplot(cmData, order = "FPC", tl.cex = 0.3, tl.col = "black")
```

```{r correlation plot 168, echo=FALSE, fig.dim=c(6, 6)}
# Select global and 16.8 m high values for correlation plot
npf168 <- data.frame(npfOrig[, (5:6)], npfOrig[, (13:16)], npfOrig[, (27:30)], 
                  npfOrig[, (41:42)], npfOrig[, (53:54)], npfOrig[, (63:72)],
                  npfOrig[, (83:88)], npfOrig[, (99:105)])
cm168 <- cor(npf168)
corrplot(cm168, order = "FPC", tl.cex = 0.5, tl.col = "black")
```

@@

```{r preprocess data, echo=FALSE}
# Sample half of the data for training and testing sets
set.seed(42)
idx <- sample(nrow(npfData), nrow(npfData)/2)
npf_train <- npfData[idx, ]
npf_test <- npfData[-idx, ]
```
